Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-18 manuscript/bef_stability/manuscript.Rmd
Decrease manuscript length to 4930

- Data accessibility statement: the ONEMA database is available upon request to 
  eddy.cosson@afbiodiversite.fr and thierry.point@afbiodiversite.fr. The code
  used for the analysis and to generate the manuscript is available on github:
  (alaindanet/fishcom.git)

@@

- Number of words:
  - Abstract: 151
  - Main text: 4980

@@

Biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships are central to ecology. However, they remain largely untested in a
foodweb context in natural settings. Here we study the links among
environmental conditions, richness, foodweb structure, annual biomass and its
temporal stability using a standardized monitoring of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that network structure affects community stability,
with a positive effect of mean trophic level, and contrary to general
expectation, we find that richness destabilizes fish communities. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in species stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006].  Species richness has
been shown to promote higher biomass production and  higher temporal stability
of total biomass in a variety of ecosystems and taxa.  Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], but
also in briophyte communities [@mulder_physical_2001], stream feeders,
[@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018].  In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020].  Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning in experimental and empirical studies remains mostly focused on
single trophic levels, ignoring the effects of foodweb structure.  This gap is
even more striking for the relationship between diversity and temporal stability
of ecosystem functioning, in spite of long-standing interest for the links
between foodweb complexity and stability in theoretical ecology [e.g.
@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

@@

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing
data from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness.

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. Despite that fish trophic interactions are
difficult to measure in natural conditions, a substantial body of literature
supports that they can be inferred through body size and diet informations
[@jennings_weak_2001; @gravel_inferring_2013], especially
in stream communities including fishes [@brose_predator_2019]. Lastly, most of
the diversity-biomass and diversity-stability relationships, have been assessed
in terrestrial ecosystems while there is a evident need to test such
relationships in a variety of ecosystem types.

Our aim is to document the relationships among species richness, foodweb
structure, total biomass and its temporal stability in natural communities. To
do so, we used the ONEMA dataset consisting in a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years spanning on the 1995-2018 period. We
inferred 1496 interaction networks, based on body size and diet. We question
whether species richness, connectance and average trophic level are related or
not to total biomass and its temporal stability. We further question whether
species richnes and foodweb structure have antagonistic or mutual effect on
total biomass and its temporal stability. Finally, we question how ecosystem size
and average temperature are related to total biomass and its temporal stability.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). The stations where the stream
was of intermediate size could be sampled alternatively with one of the two
sampling protocols. In this case, we kept only the most frequently used sampling
protocol. The sampling season spanned from late spring to autumn. There was low
heterogeneity of the sampling month within stations, with the median sampled
month being mid-August while the median of the standard deviation of the sampled
month within station was $0.7$.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in total biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet) A).

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Species biomass and total biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, species biomass and
total biomass were corrected by the surface sampled (range `r range_surface[1]`-
`r range_surface[2]` square metres), and expressed in grams by square meter
($gm^{-2}$). In the analysis, the total biomass of a site has been defined as
the median of the total biomass across sampling years.

@@

We described the mechanisms driving the temporal stability of biomass, by
decomposing the $CV$ of total biomass ($CV_{com}$) into a synchrony ($\phi$) and
a weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

Species richness in a given station has been defined as the total number of species present across
all the sampling events. We controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species per square
meter sampled.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Fig.
\@ref(fig:matmet) B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Fig. \@ref(fig:matmet) C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet) B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

@@

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic specie was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Fig. S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class, for both the piscivorous and non piscivorous trophic species.
Lastly, the trophic links among resource nodes were set according to the
literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed two of the most commonly used network metrics to characterize food
webs, namely the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the maximum number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of trophic
species plus resource nodes). The mean trophic level $\overline{T}$ was computed
as the mean trophic level weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and
$B_i$ being respectively the trophic level and the biomass of the node $i$. The
trophic level of the nodes was computed as one added to the average trophic
level of its food item. For each site, the median of connectance and average
trophic level over time was used in subsequent analysis.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow. As data from the naiades database are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, DBO and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure as measured by connectance and
average trophic level [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Fig.
\@ref(fig:sem-plot) A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Fig.
\@ref(fig:sem-plot) C). Instead of the annual total biomass, we modelled the
synchrony and population variability which in turn determined the temporal
stability of biomass.

@@

Indirect effects on total biomass and its temporal stability
were computed by multiplying slopes along the path. As an example, if we consider
$A \to B \to C$, the indirect effect of A on C is equal to $r_{AB} \times
r_{BC}$. Only significant direct slopes were entered in the calculation of
indirect effects. Total effects were computed as the sum
of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and the manuscript are accessible on github (alaindanet/fishcom). The analyses were
performed with R version 3.6.3 [@r_core_team_r_2020].

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 =
.11$, Fig. \@ref(fig:sem-plot) B and D). Species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_ct`$, Fig.
\@ref(fig:sem-plot) B and D), meaning that a part of the effects of species
richness went through network structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-plot) B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect on total biomass
(Fig. \@ref(fig:sem-plot) B).

The two PCA axes capturing environmental variables had no significant direct
effects on total biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, PCA1, related to the stream size was positively linked to total biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, PCA2, related to the average
annual temperature and inversely to altitude of the sites was positively related
to total biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). The total effect
of PCA1 and PCA2 on total biomass was of the same magnitude than the total effects of species
richness on total biomass.

@@

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-plot) D).

The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-plot) D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)).

>>>>

- Data accessibility statement: Upon acceptance of the paper, we confirmed that
  the data supporting the results will be archived in an appropriate public
  repository and that the data DOI will be included at the
  end of the article. The code used for the analysis and to generate the
  manuscript is available on github: (alaindanet/fishcom.git). The raw fish
  database is public and available upon request to eddy.cosson@afbiodiversite.fr and
  thierry.point@afbiodiversite.fr. The raw data of water temperature, flow and
  Biological Oxygen Demand are available on http://www.naiades.eaufrance.fr/.

@@

- Number of words:
  - Abstract: 149
  - Main text: 4930

@@

Biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships are central to ecology. However, they remain largely untested for 
foodweb in natural context. Here we study the links among
environmental conditions, richness, foodweb structure, annual biomass and its
temporal stability using a standardized monitoring of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that network structure affects community stability,
with a positive effect of mean trophic level, and contrary to general
expectation, we find that richness destabilizes fish communities. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in species stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006]. Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], stream
feeders, [@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018]. In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020]. Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning remains mostly focused on
single trophic levels, ignoring the effects of foodweb structure. This gap is
even more striking for the relationship between diversity and temporal stability
, in spite of long-standing interest for the links
between foodweb complexity and stability in theoretical ecology [e.g.
@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

@@

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural
communities, the relationships among species richness, total biomass and its
temporal stability might also arise from effects of the environment on both
diversity and ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016].

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @gravel_inferring_2013], especially in stream
communities including fishes [@brose_predator_2019].

Our aim is to document the relationships among species richness, foodweb
structure, total biomass and its temporal stability in natural communities. To
do so, we used a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years. We inferred 1496 interaction
networks, based on body size and diet. We question whether species richness,
connectance and average trophic level are related to total biomass and its
temporal stability. We further question how ecosystem size and temperature are
related to total biomass and its temporal stability.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electrofishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). In the case where a stream was
sampled alternatively with one of the two sampling protocols, we kept only the
most frequently used. The sampling season spanned from late spring to autumn.
There was low heterogeneity of the sampling month within stations, with the
median sampled month being mid-August while the median of the standard deviation
of the sampled month within station was $0.7$.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in total biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Species biomass and total biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, species biomass and
total biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the total biomass of a site has been defined as
the median of the total biomass across sampling years.

@@

We described the mechanisms driving the temporal stability of biomass,
decomposing the $CV$ of total biomass ($CV_{com}$) into a synchrony ($\phi$) and
a weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the CV of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average CV of
species in the station $i$ weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

Species richness in a given site was defined as the total number of species present across
all the sampling events. We controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species per square
meter sampled.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Fig.
\@ref(fig:matmet)B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

@@

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Fig. S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class.  Lastly, the trophic links among resource nodes were set according
to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of nodes).
The trophic level of the nodes was computed as one added to the average trophic
level of its food item. The mean trophic level $\overline{T}$ was 
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i \times
(\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the trophic
level and the biomass of the node $i$. For each site, the median of connectance
and average trophic level over time was used in subsequent analysis.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. We obtained data on water temperature and flow from the naiades
database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Fig. \@ref(fig:sem-plot)A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Fig.
\@ref(fig:sem-plot)C). Instead of the annual total biomass, we modelled the
synchrony and $\overline{CV_{sp}}$ which in turn determined the temporal
stability of biomass.

@@

Indirect effects on total biomass and its temporal stability
were computed by multiplying slopes along the path. Only significant direct
slopes were entered in the calculation of indirect effects. Total effects were
computed as the sum of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and to generate the manuscript are accessible on github
(alaindanet/fishcom). The analyses were performed with R version 3.6.3
[@r_core_team_r_2020].

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 =
.11$, Fig. \@ref(fig:sem-plot)B and D). Species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_ct`$, Fig.
\@ref(fig:sem-plot)B and D), meaning that a part of the effects of species
richness went through network structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-plot)B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect.

The two PCA axes capturing environmental variables had no significant direct
effects on total biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, related to the stream size, was positively linked to total biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, related to the average
annual temperature was positively related
to total biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). The total effect
of PCA1, PCA2 and species richness on total biomass were of the same magnitude.

@@

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-plot)D).

The total effect of mean trophic level and species richness on temporal
stability of biomass were of the same magnitude (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-plot)D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-plot)D, Table \@ref(tab:indir-sem)).

The total effect of PCA2, related to average annual temperature, on temporal
stability was of higher magnitude than the effects of species richness and mean
trophic level (Table \@ref(tab:indir-sem)).
The PCA2 had a total negative effect on temporal stability, meaning that higher
temperature results in lower stability
($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table \@ref(tab:indir-sem)). 
PCA2 had an indirect positive effect on temporal stability through a
direct positive effect on mean trophic level ($r_\delta = 
`r env_on_stab$via_tlvl["log_RC2"]`$, Table \@ref(tab:indir-sem)). However, PCA2
had a negative effect on temporal stability through a direct positive effect on 
$\overline{CV_{sp}}$ and species richness (resp.  $r_\delta = 
`r env_on_cv_sp$direct["log_RC2"]`$ and $r_\delta = `r env_on_stab$via_richness["log_RC2"]`$,
Table \@ref(tab:indir-sem)). This was also true for the PCA1, related to stream
size, with a resulting very slight positive effect of PCA1 on temporal stability
($r_\delta = `r env_on_stab$total["log_RC1"]`$, Table \@ref(tab:indir-sem)).
